---
title: "Making maps of Ireland in R"
author: "Brendan O'Dowd"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This guide uses maps from my GitHub page produced using shapefiles from [OSi](https://data-osi.opendata.arcgis.com/) and [OSNI](https://www.spatialni.gov.uk/).

Two packages are used throughout, they are [tidyverse](https://tidyverse.tidyverse.org/) (actually a collection of packages) and [sf](https://r-spatial.github.io/sf/), which stands for simple features. To install any package run ```install.packages("package_name")```.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
