---
title: "Making maps of Ireland in R"
author: "Brendan O'Dowd"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This guide uses maps from my GitHub page produced using shapefiles from [OSi](https://data-osi.opendata.arcgis.com/) and [OSNI](https://www.spatialni.gov.uk/).
 
Two packages are used throughout, they are [tidyverse](https://tidyverse.tidyverse.org/) (actually a collection of packages) and [sf](https://r-spatial.github.io/sf/), which stands for simple features. To install any package run ```install.packages("package_name")```.

To start, we will need to load both of these packages

```{r packages , message=FALSE}
library(tidyverse)
library(sf)
```

## Importing maps

You **should** be able to import my maps directly from the web, using the following, which will create an object called ```lea_166```. This is an sf object, it is a special type of dataframe that has a column called geometry which contains the coordinates for the shapes. This is a map of the 166 Local Electoral Areas in Ireland, plus the outline for Northern Ireland, so there are 167 shapes in total and 167 rows in ```lea_166```

You can also embed plots, for example:

```{r import_map , message=FALSE}
lea_166 <- st_read("https://raw.githubusercontent.com/brendanjodowd/maps/main/lea_166.geojson")
```

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
